Mixed-kNN Comparison (fvc)
==========================

```{r echo=FALSE, results="hide"}
opts_chunk$set(tidy = FALSE)
library(ggplot2)

for (fn in dir("../../R/", full.names = TRUE)) source(fn)

data <- ReadData("../../data/train.csv")
train <- subset(data, year < 1.5)
train.pids <- unique(train[["PtID"]])
test <- subset(data, year >= 1.5 & PtID %in% train.pids)

num.train.visits <- ddply(data, .(PtID), summarize, nvisits = sum(year < 1.5))
```

First, fit the mixed effects predictor model.

```{r mem, cache=TRUE}
mem.model <- MixedModelPredictor(train, "year", "fvc")
mem.results <- predict(mem.model, test)
```

Fit the kNN predictor model.

```{r knn, cache=TRUE}
knn.model <- KnnPredictor(train, k = 5)
knn.results <- predict(knn.model, test)
```

Performance
-----------

Compute the error measurements.

```{r}
if (nrow(mem.results) != nrow(knn.results))
    stop("Different number of predictions")

mem.errors <- ddply(mem.results, .(PtID),
                    summarize, norm.mse = mean(((fvc.hat - fvc) / fvc.sd) ^ 2))

knn.errors <- ddply(knn.results, .(PtID),
                    summarize, norm.mse = mean(((fvc.hat - fvc) / fvc.sd) ^ 2))

```

Plot the mean errors.

```{r}
mem.errors[["method"]] <- "mem"
knn.errors[["method"]] <- "knn"
errors <- rbind(mem.errors, knn.errors)

summary(subset(errors, method == "mem")[["norm.mse"]])
summary(subset(errors, method == "knn")[["norm.mse"]])

p <- ggplot(subset(errors, norm.mse < 2), aes(x = norm.mse))
p + geom_histogram() + facet_wrap(~ method)
```

Plot the errors for predictions less than 4.5 years from baseline.

```{r}
mem.short.errors <- ddply(subset(mem.results, year < 4.5), .(PtID),
                          summarize, norm.mse = mean(((fvc.hat - fvc) / fvc.sd) ^ 2))

knn.short.errors <- ddply(subset(knn.results, year < 4.5), .(PtID),
                          summarize, norm.mse = mean(((fvc.hat - fvc) / fvc.sd) ^ 2))

mem.short.errors[["method"]] <- "mem"
knn.short.errors[["method"]] <- "knn"
short.errors <- rbind(mem.short.errors, knn.short.errors)

summary(subset(short.errors, method == "mem")[["norm.mse"]])
summary(subset(short.errors, method == "knn")[["norm.mse"]])

p <- ggplot(short.errors, aes(x = norm.mse))
p + geom_histogram() + facet_wrap(~ method)
```

Plot the errors excluding patients who have less than 2 training visits.

```{r}
at.least.two <- unique(subset(num.train.visits, nvisits >= 2)[["PtID"]])

mem.e <- subset(mem.errors, PtID %in% at.least.two)
mem.e[["method"]] <- "mem"

summary(mem.e[["norm.mse"]])

knn.e <- subset(knn.errors, PtID %in% at.least.two)
knn.e[["method"]] <- "knn"

summary(knn.e[["norm.mse"]])

pruned.errors <- rbind(mem.e, knn.e)

p <- ggplot(short.errors, aes(x = norm.mse))
p + geom_histogram() + facet_wrap(~ method)
```

Analysis (mem)
--------------

Examine some patients who have large norm.mse under the mixed effects model.

```{r}
patients <- arrange(subset(mem.errors, norm.mse < 2), desc(norm.mse))[1:5, ]
print(patients)

patient.data <- subset(data, PtID %in% patients[["PtID"]])
patient.coef <- subset(coef(mem.model), PtID %in% patients[["PtID"]])
patient.knn.pred <- subset(knn.results, PtID %in% patients[["PtID"]])

p <- ggplot(patient.data, aes(x = year, y = fvc, color = PtID, shape = factor("true")))
p + geom_point(size = 2) +
    geom_point(aes(x = year, y = fvc.hat, color = PtID, size = 2, shape = factor("knn")), patient.knn.pred) +
    geom_abline(aes(intercept = intercept, slope = slope, color = PtID),
                patient.coef)
```

For the five patients we've examined above, plot their nearest neighbors.

```{r}
patient.ids <- unique(patients[["PtID"]])

PlotNeighbors <- function(pid) {
    patient.data <- subset(data, PtID == pid)
    
    k.nearest <- coef(knn.model, pid)
    print(k.nearest)
    k.nearest.data <- subset(data, PtID %in% k.nearest)

    stopifnot(nrow(k.nearest.data) > 0)

    ggplot(patient.data) +
        geom_line(aes(x = year, y = fvc, color = factor(PtID))) +
        geom_point(aes(x = year, y = fvc, color = factor(PtID)), data = k.nearest.data)
}

PlotNeighbors(patient.ids[1])
PlotNeighbors(patient.ids[2])
PlotNeighbors(patient.ids[3])
PlotNeighbors(patient.ids[4])
PlotNeighbors(patient.ids[5])
```

For comparison, produce a similar plot for patients who are predicted well.

```{r}
patients <- arrange(mem.errors, norm.mse)[1:5, ]
print(patients)

patient.data <- subset(data, PtID %in% patients[["PtID"]])
patient.coef <- subset(coef(mem.model), PtID %in% patients[["PtID"]])

p <- ggplot(patient.data, aes(x = year, y = fvc, color = PtID))
p + geom_point() +
    geom_abline(aes(intercept = intercept, slope = slope, color = PtID),
                patient.coef)
```

Analysis (knn)
--------------

Examine patients who have large norm.mse under k-NN.

```{r}
patients <- arrange(knn.errors, desc(norm.mse))[1:5, ]
print(patients)

patient.data <- subset(data, PtID %in% patients[["PtID"]])
patient.pred <- subset(knn.results, PtID %in% patients[["PtID"]])

ggplot() +
    geom_point(aes(x = year, y = fvc, color = PtID, shape = factor("true")), patient.data) +
    geom_point(aes(x = year, y = fvc.hat, color = PtID, shape = factor("predicted")), patient.pred)
```

Similar plot for patients who are predicted well.

```{r}
patients <- arrange(knn.errors, norm.mse)[1:5, ]
print(patients)

patient.data <- subset(data, PtID %in% patients[["PtID"]])
patient.pred <- subset(knn.results, PtID %in% patients[["PtID"]])

ggplot() +
    geom_point(aes(x = year, y = fvc, color = PtID, shape = factor("true")), patient.data) +
    geom_point(aes(x = year, y = fvc.hat, color = PtID, shape = factor("predicted")), patient.pred)
```

Mixed-kNN Comparison (dlco)
===========================
