Variable Plots
==============

```{r echo=FALSE}
opts_chunk$set(fig.align = "center")
```

```{r}
library(ggplot2)
library(plyr)
library(reshape2)
source("../../R/read_data.R")

data <- ReadData("../../data/fvc.5.visits.csv")
```

What is the distribution over FVC and DLCO.

```{r}
p <- ggplot(data, aes(x = year, y = fvc))
p + geom_point(alpha = 0.1)

p <- ggplot(data, aes(x = year, y = dlco))
p + geom_point(alpha = 0.1)
```

How do FVC and DLCO vary together. It's interesting that they correlations are not always close to 1 for every patient.

```{r}
data.summary <- ddply(data, .(PtID), summarize,
                      fvc.dlco.cov = cov(fvc, dlco),
                      fvc.dlco.cor = cor(fvc, dlco))

p <- ggplot(data.summary, aes(x = fvc.dlco.cor))
p + geom_histogram()

```

Examine some of the patients where FVC and DLCO are not correlated.

```{r}
pids <- arrange(data.summary, fvc.dlco.cor)$PtID[1:5]
patient.data <- subset(data, PtID %in% pids, select = c(PtID, year, fvc, dlco))
patient.data <- melt(patient.data, measure.vars = c("fvc", "dlco"))

p <- ggplot(patient.data,
            aes(x = year,
                y = value,
                group = interaction(PtID, variable),
                color = factor(PtID)))
p + geom_line()
```

How does FVC change when interacting with antibodies.

```{r}
p <- ggplot(data, aes(x = fvc, fill = interaction(aca, topo, rnp, pol3)))
p + geom_density(alpha = 0.2)
```

What is the distribution over FVC when conditioning on the number of
visits that a patient has.

```{r}
df <- ddply(data, .(PtID), transform, num.visits = length(fvc))
p <- ggplot(df, aes(x = factor(num.visits), y = fvc))
p + geom_boxplot()
```

What is the best way to characterize a patient's overall FVC level?
First, we'll try to take the top 10% of the data for each patient and
report the median of that subset.

```{r}
df <- ddply(data, .(PtID), summarize, x = median(fvc[fvc > quantile(fvc, 0.9)]))
qplot(x, data=df)
```

Can we use this statistic to see how well FVC and DLCO correlate?

```{r}
ComputeQuantMedian <- function(x) median(x[x > quantile(x, 0.9)])
df <- ddply(data, .(PtID), summarize,
            fvc.s = ComputeQuantMedian(fvc),
            dlco.s = ComputeQuantMedian(dlco))
qplot(scale(fvc.s), scale(dlco.s), data = df)
```
