Measurements
============

```{r}
opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 7, dev = "svg")

visit.filename <- "../../../data/sclerodata/tVisit.csv"
pft.filename <- "../../../data/sclerodata/tPFT.csv"
```

Consecutive Measurement Variation
---------------------------------

The following four plots show the average difference between
measurements that are taken within some number of years of one
another. The title of each plot displays which measurement is being
shown. The number of years is swept from 0.5 to 3 years by half year
steps.

**Takeaway**: At 1.5, 2.0, 2.5, and 3.0 year thresholds, the histograms
 are slightly skewed to the left. I think that this reflects the
 population-level downward trend in the data. With this in mind, I
 think that the true variability of each of the types of measurements
 falls within the range seen in the 0.5 and 1 year threshold plots.

```{r}
source("../lab_var_sequential_diff.R")
```

```{r fvc-variation}
main(pft.filename, "FVC.Pre")
```

```{r perc-fvc-variation}
main(pft.filename, "perc.FVC.of.predicted")
```

```{r dlco-variation}
main(pft.filename, "DLCO")
```

```{r perc-dlco-variation}
main(pft.filename, "perc.DLCO.of.predicted")
```

Severity Crossplots
-------------------

Each of the lab and clinical measurements captures some aspect of the
disease's severity for a patient. We are interested in understanding
the extent to which these different measurements are correlated or
uncorrelated. For example, we want to know whether a low GI severity
score correlates with a low FVC measurement.

```{r}
source("../severity_crossplot.R")
sev.data <- main(visit.filename, pft.filename)
```

This figure shows the interactions between categorical severity
scores, which are recorded during clinical visits.

The first thing to note is that the skin severity score can increase
without the Raynaud's severity score increasing. Raynaud's measures
the condition of the skin on a patient's digits, so it makes sense
that the skin score severity (which measures the condition of the skin
across the patient's entire body) can become worse without RP
increasing in severity as well.

This also seems to be the case for GI severity, since it seems that
the GI score can become quite severe while the RP score remains at 1
or 2.

It is interesting, though, that skin severity and GI severity do not
seem to be correlated. The top right subplot shows that the number of
patients who have both high skin and high GI scores is relatively
small.

```{r clinic-var-cross}
plot_categorical(sev.data)
```

This plot does not demonstrate anything too surprising. We see that
DLCO and FVC.Pre correlated.

```{r lab-var-cross}
plot_continuous(sev.data)
```

This plot summarizes the interactions between FVC and the 3 clinical
measurements. The top three subplots on the right are the most
important, and show the distribution over FVC severity for each value
of the 3 clinical measurements. It seems that both skin severity and
GI severity are correlated with a decrease in FVC. RP severity, on the
other hand, does not seem to have as strong of a correlation. Although
patients with an RP severity score of 0 have better FVCs than patients
with a score of 1, 2, 3, or 4, the severity of FVC does not seem to
get worst as RP becomes more severe.

```{r fvc-clinic-cross}
plot_combo_fvc(sev.data)
```

DLCO interacts with skin severity and GI severity as FVC does, but
there seems to be a slightly stronger correlation with RP
severity. Although DLCO and FVC are correlated in general, there are
some outliers. I'd like to look into whether there is an overlap with
these outliers and the patients who have a lower DLCO and a higher RP
severity score.

```{r dlco-clinic-cross}
plot_combo_dlco(sev.data)
```
