Mixed Effects Prediction 2
==========================

```{r results="hide"}
opts_chunk$set(fig.align="center", tidy=FALSE)
library(ggplot2)
library(lme4)
library(plyr)

source("../../R/read_data.R")
source("../../R/plot_results.R")
```

Only use year, gender, and race to predict for now.

```{r}
data <- ReadData("../../data/fvc.5.visits.csv")
```

Fit a simple mixed effects model using the first 1.5 years of data.

```{r}
train <- subset(data, year < 1.5)
mem.fit <- lmer(fvc ~ year + strFemale + strRaceId1 + (1 + year|PtID),
                data = train)
print(mem.fit)
```

Check how well the model fits the training data. The residuals don't seem to have any systematic bias, so the model fits the first 1.5 years of data relatively well.

```{r fig.align="center"}
residuals <- data.frame(year = train$year,
                        fvc.sd = train$fvc.sd,
                        residual = resid(mem.fit))

p <- ggplot(residuals, aes(x = year, y = residual / fvc.sd))
p + geom_point(alpha = 0.3) + geom_smooth()
```

Predict the rest of the data for each patient using the fixed and random effects fit above.

```{r}
patient.coef <- coef(mem.fit)$PtID

factor.names <- names(patient.coef)[3:length(patient.coef)]
names(patient.coef) <- c("intercept", "slope", factor.names)
patient.coef$PtID <- rownames(patient.coef)

test <- merge(subset(data, year >= 1.5), patient.coef, by = "PtID")

test$fvc.hat <- with(test, {
    intercept + slope * year +
        ifelse(strFemale == 0, 0.0, strFemale1) +
        ifelse(strRaceId1 == 0, 0.0, eval(as.name(paste0("strRaceId1", strRaceId1))))
})

test$resid <- test$fvc.hat - test$fvc
mean(test$resid ** 2)
```

Plot the predicted residuals.

```{r}
p <- ggplot(test, aes(x = year, y = resid / fvc.sd))
p + geom_point(alpha = 0.3) + geom_smooth()
```

Plot the patients with high MSE.

```{r}
source("../../R/plot_results.R")
PlotExtremes(data, test, extreme = "high", k = 5)
```

Plot the patients with low MSE.

```{r}
PlotExtremes(data, test, extreme = "low", k = 5)
```